### 2021-04-15 com/c1.tex
WIP 16Apr2021

Let us now examine the converse. Let $T(x_i) = \lambda_i x_i$. Introduce the
variables
\begin{eqnarray*}
A_{ij} &=& \alpha_i\alpha_j\ip{x_i}{x_j} \\
B_{ij} &=& \beta_i\beta_j\ip{x_i}{x_j} \\
C_{ij} &=& \alpha_i\beta_j\ip{x_i}{x_j}
\end{eqnarray*}
so that
\begin{eqnarray*}
\angle(a, b) &=& \frac{\sum_{r,s}C_{rs}}
                 {\sqrt{\sum_{k,l}A_{kl}}\sqrt{\sum_{m,n}B_{mn}}} \\
(\angle(a, b))^2 &=& \frac{\sum_{rstu}C_{rs}C_{tu}}
                 {\sum_{k,l}A_{kl}\sum_{m,n}B_{mn}}  \\
(\angle(T(a), T(b)))^2 &=& \frac{\sum_{rstu}\lambda_r\lambda_s\lambda_t\lambda_u
    C_{rs}C_{tu}}
    {\sum_{k,l}\lambda_k\lambda_l A_{kl}\sum_{m,n}\lambda_m\lambda_n B_{mn}} 
\end{eqnarray*}
When $T$ is angle preserving,
\[
\sum_{k,l,m,n,r,s,t,u}\lambda_k\lambda_l\lambda_m\lambda_n A_{kl}B_{mn}C_{rs}
C_{tu} = 
\sum_{k,l,m,n,r,s,t,u}\lambda_r\lambda_s\lambda_t\lambda_u A_{kl}B_{mn}C_{rs}
C_{tu}
\]
or
\[
\sum_{k,l,m,n,r,s,t,u}(\lambda_k\lambda_l\lambda_m\lambda_n - 
\lambda_r\lambda_s\lambda_t\lambda_u)A_{kl}B_{mn}C_{rs}C_{tu} = 0.
\]
This equation is true for all vectors $a, b$, that is all possible values of 
$A_{kl}, B_{mn}, C_{rs}$ and $C_{tu}$. Therefore, we must have
\[
\lambda_k\lambda_l\lambda_m\lambda_n = \lambda_r\lambda_s\lambda_t\lambda_u
\]
for all $k, l, m, n, r, s, t, u$. In particular, it is true for $k = l = m = n$
and $r = s = t = u$ so that $\lambda_k^4 = \lambda_r^4$ or $\lambda_k = \pm
\lambda_r$.
\end{enumerate}
\end{enumerate}

@@

\section{Subsets of euclidean sets}\label{c1s2}
\begin{thm}[Heine-Borel]
$[a, b]$ is compact.
\end{thm}
\begin{proof}
Let $\mathcal{O}$ be an open cover of $[a, b]$ and define the set
\[
X = \{x : [a, x] \text{ has a finite open cover}\}.
\]
The set $X$ is non-empty because $a \in X$. It is also bounded. Therefore, it
has a supremum, say $L$.

@@

>>>>

Let us now examine the converse. Let $T(x_i) = \lambda_i x_i$. Introduce the
variables
\begin{eqnarray*}
A_{ij} &=& \alpha_i\alpha_j\ip{x_i}{x_j} \\
B_{ij} &=& \beta_i\beta_j\ip{x_i}{x_j} \\
C_{ij} &=& \alpha_i\beta_j\ip{x_i}{x_j}
\end{eqnarray*}
so that
\begin{eqnarray*}
\angle(a, b) &=& \frac{\sum_{r,s}C_{rs}}
                 {\sqrt{\sum_{k,l}A_{kl}}\sqrt{\sum_{m,n}B_{mn}}} \\
(\angle(a, b))^2 &=& \frac{\sum_{rstu}C_{rs}C_{tu}}
                 {\sum_{k,l}A_{kl}\sum_{m,n}B_{mn}}  \\
(\angle(T(a), T(b)))^2 &=& \frac{\sum_{rstu}\lambda_r\lambda_s\lambda_t\lambda_u
    C_{rs}C_{tu}}
    {\sum_{k,l}\lambda_k\lambda_l A_{kl}\sum_{m,n}\lambda_m\lambda_n B_{mn}} 
\end{eqnarray*}
When $T$ is angle preserving,
\[
\sum_{k,l,m,n,r,s,t,u}\lambda_k\lambda_l\lambda_m\lambda_n A_{kl}B_{mn}C_{rs}
C_{tu} = 
\sum_{k,l,m,n,r,s,t,u}\lambda_r\lambda_s\lambda_t\lambda_u A_{kl}B_{mn}C_{rs}
C_{tu}
\]
or
\[
\sum_{k,l,m,n,r,s,t,u}(\lambda_k\lambda_l\lambda_m\lambda_n - 
\lambda_r\lambda_s\lambda_t\lambda_u)A_{kl}B_{mn}C_{rs}C_{tu} = 0.
\]
This equation is true for all vectors $a, b$, that is all possible values of 
$A_{kl}, B_{mn}, C_{rs}$ and $C_{tu}$. Therefore, we must have
\[
\lambda_k\lambda_l\lambda_m\lambda_n = \lambda_r\lambda_s\lambda_t\lambda_u
\]
for all $k, l, m, n, r, s, t, u$. In particular, it is true for $k = l = m = n$
and $r = s = t = u$ so that $\lambda_k^4 = \lambda_r^4$ or $\lambda_k = \pm
\lambda_r$.
\end{enumerate}

\item The transformation
\[
T = \begin{bmatrix}\cos\theta&\sin\theta\\-\sin\theta&\cos\theta\end{bmatrix}
\]
is a rotation by an angle $\theta$ therefore it is angle-preserving. 
Nevertheless, one can formally confirm its nature by checking if it is norm-
preserving. Consider 
\[
y = T(x) = \begin{bmatrix}x^1\cos\theta + x^2\sin\theta \\ 
-x^1\sin\theta + x^2\cos\theta\end{bmatrix}
\]
so that
\begin{eqnarray*}
\norm{y} &=& (x^1)^2\cos^2\theta + (x^2)^2\sin^2\theta + 2x^1x^2\sin\theta
\cos\theta \\
 & & + (x^1)^2\sin^2\theta + (x^2)^2\cos^2\theta - 2x^1x^2\sin\theta
\cos\theta \\
 &=& (x^1)^2 + (x^2)^2 \\
 &=&  \norm{x}.
\end{eqnarray*}

\item $T$ is a fixed linear transformation. Let $h = (h^1, \ldots, h^m)$,
then the $j$th component of $T(h)$ is $T(h)^j = T_{ij}h^j$ and the norm of
$T(h)$ is
\[
\norm{T(h)} = \sum_{j=1}^m (T_{ij}h^j)^2
\]
Let $M$ be the largest among $\abs{T_{ij}}$ so that
\[
\norm{T(h)} \le M \sum_{j=1}^m (h^j)^2 = M\norm{h}.
\]
\begin{rem} This exercise shows that a linear transformation does not lead to
a blow up.
\end{rem}

\item Quite straightforward.

\item $T : \sor^n \rightarrow (\sor^n)^\ast$ is such that $T(x) = \varphi_x$.
Consider $T(x_1 + x_2) = \varphi_{x_1 + x_2}$ so that
\begin{eqnarray*}
T(x_1 + x_2)(y) &=& \varphi_{x_1+x_2}(y) \\
 &=& \ip{x_1+x_2}{y} \\
 &=& \ip{x_1}{y} + \ip{x_2}{y} \\
 &=& \varphi_{x_1}(y) + \varphi_{x_2}(y) \\
 &=& T(x_1)(y) + T(x_2)(y) \\
 &=& (T(x_1) + T(x_2))(y)
\end{eqnarray*}
\begin{rem}
$\varphi_x$ is called a linear functional. The space of all linear functionals,
the dual space $(\sor^n)^\ast$ is a linear space so is the set of 
transformations $T(x) = \varphi_x$.
\end{rem}

\item Follows immediately from the fact that $\norm{x+y}^2 = \ip{x+y}{x+y}$ 
and the linearity of the inner product.
\end{enumerate}

@@

\section{Subsets of euclidean sets}\label{c1s2}
\begin{lem}\label{c1s2l1}
Every set $X$ has an open cover.
\end{lem}
\begin{proof}
$\{\cup_x B_\epsilon(x) : x \in X, \epsilon > 0\}$ is an open cover, where 
$B_\epsilon(x)$ is an open ball of radius $\epsilon$ centred at $x$.
\end{proof}

\begin{thm}[Heine-Borel]\label{c1s2t1}
$[a, b]$ is compact.
\end{thm}
\begin{proof}
Let $\mathcal{O}$ be an open cover of $[a, b]$ and define the set
\[
X = \{x : [a, x] \text{ has a finite open cover}\}.
\]
The set $X$ is non-empty because $a \in X$. It is also bounded. Therefore, it
has a supremum, say $L$.

@@

\begin{lem}\label{c1s2l2}
If $B \subset \sor^m$ is a compact set then for any $x \in \sor^n$, the set
$\{x\} \times B$ is compact.
\end{lem}
\begin{proof}
The finite open cover $A_1, \ldots, A_k$ of $B$ can be extended to $\{x\} \times
A_1, \ldots, \{x\} \times A_k$, which is a finite open cover of $\{x\}\times B$.
\end{proof}

\begin{thm}\label{c1s2t2}
If $B \subset \sor^m$ is a compact set and $\mathcal{O}$ is an open cover of
$\{x\} \times B$, for an $x \in \sor^n$, then there exists a subset $U$ of
$\sor^n$, containing $x$, such that $U \times B$ is covered by a finite number
of members of $\mathcal{O}$.
\end{thm}
\begin{proof}
From \eqref{c1s2l1} we can assume that $\mathcal{O}$ is a finite set. Consider
$\{x\} \times B$. Let $\{x\} \times W_1, \ldots \{x\} \times W_k$ be its cover.
Each one of these has an open rectangle $Q_1 \times R_1, \ldots R_k \times Q_k$
such that $Q_1, \ldots Q_k$ are subsets of $\sor^n$ and $R_1, \ldots, R_k$ are
subsets of $\sor^m$. Let $U = Q_1 \cap \cdots \cap Q_k$. Then $U \times B$ is
covered by $Q_1 \times R_1, \ldots, Q_k \times R_k$.
\end{proof}

This theorem assures that one can always extend the finite open cover of a 
subset of $\sor^m$ to an open cover of a subset of $\sor^n \times \sor^m$.

\begin{prop}\label{c1s2p1}
If $A \subset \sor^n$ and $B \subset \sor^m$ are compact then $A \times B 
\subset \sor^n \times \sor^m$ is compact.
\end{prop}
\begin{proof}
If $U_1, \ldots, U_k$ is an open cover of $A$ and $V_1, \ldots, V_l$ that of
$B$ then the sets $U_i \times V_j$ where $i = 1, \ldots, k$ and $j = 1, \ldots,
l$ is a finite open cover of $A \times B$.
\end{proof}

As a result of theorem \ref{c1s2t1} and proposition \ref{c1s2p1} we have
\begin{lem}\label{c1s2l3}
A closed bounded set of $\sor^n$ is compact.
\end{lem}

\subsection{Some additional lemmas}
\begin{lem}\label{c1s2l4}
A set if open if every element of it lies in a open ball entirely in the set.
\end{lem}
\begin{proof}
By definition, a set is open if every element of it lies in an open hyper-
rectangle entirely in the set. Choose the rectangle to be a hypercube of side 
$a$ centered at the element. Then we can enclose a ball of radius $a/2$ with 
centre coinciding the hypercube's centre.
\end{proof}

\subsection{Problems}
\begin{enumerate}
\item It is easy to check that a union, countable or not, of open sets is open.
Now consider a finite intersection of open sets. If $x \in U_1, \ldots, U_k$
then there are open balls $B^1_{\epsilon_1}(x), \ldots, B^k_{\epsilon_k}(x)$
entirely in $U_1, \ldots, U_k$. The ball $B_\epsilon(x)$ where $\epsilon = 
\min(\epsilon_1, \ldots, \epsilon_k)$ lies entirely in $U_1 \cap \cdots U_k$,
making it an open set.

Infinite unions may not be open because the infimum of $\epsilon_1, \epsilon_2,
\ldots$ may be zero leaving us with a singleton set after taking an intersection
and a singleton set is always closed.

As an example consider the open sets is $U_n = (-1/n, 1/n)$, $n \in \soi^+$.
Their intersection is $\{0\}$.

\begin{rem}
That is why a topological space is defined as the pair $(X, \tau)$, where $\tau$
is a collection of subsets of $X$ such that
\begin{enumerate}
\item $X$ and $\varnothing$ belong to $\tau$.
\item An arbitrary union of sets of $\tau$ belong to $\tau$.
\item A \underline{finite} intersection of sets of $\tau$ belong to $\tau$.
\end{enumerate}
We want a topological space to be closed under operations on open sets.
\end{rem}

\item Proved as lemma \ref{c1s2l4}.

\item 
\begin{enumerate}
\item[(a)] $U = \{x \in \sor^n : \norm{x} \le 1\}$. The interior is the open
ball $B_1(0)$, the boundary is the sphere $S^2$ and the exterior is everything
outside it.
\item[(b)] $S^2 = \{x \in \sor^n : \norm{x} = 1\}$. There is no interior. 
Neither is there any exterior. $S^2$ is its own boundary.
\item[(c)] $U = \{x \in \sor^n : x^i \in \soq\}$. There is no neighbourhood of 
$x \in U$ with points only in $U$. The same is true for the set $\sor^n - U$.
Therefore, $U$ is neither open nor closed. It has no interior or exterior and
it is its own boundary.
\end{enumerate}

\item If $A$ was defined as $(0, 1) \times (0, 1)$ and there was no mention of
the boundary then a diagonal line would have met the restrictions of the 
problem.

If $A = [0, 1] \times [0, 1]$ then let $i_n, j_n = 2^{-n}$ for all $n \in \son$.
Then the points $H=\{0, i_n, 1 - i_n, 1\}$ divide $[0, 1]$ in $n$ intervals of
equal length. Same is true for the points $V = \{0, j_n, 1 - j_n, 1\}$. The
set $H \times V$ defines a grid of equally sized cells of the closed rectangle
 $A$. Choose one point from each of these cells.

\item $A$ is an open set and is a subset of $[0, 1]$. The boundary of $A$
contains points all neighbourhoods of which contain some points in $A$ and
some in its exterior. If $E$ is the exterior of $A$ and $\partial A$ is its
boundary then $E \cap \partial A = [0, 1] - A$.

It suffices to show that $E$ is empty. Let $x \in E$. Then there exists a
neighbourbood $(x - \epsilon, x + \epsilon)$ such that there is no point of
$A$ in it. But that is impossible because there is at least one rational number
in this neighbourhood. Thus, $E = \varnothing$ and hence $\partial A = [0, 1]
- A$.

\end{enumerate}
